---
title: "p8105_hw2_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-10-06"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

## NYC Subway Data

```{r transit}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data - NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```
### Subway Data Description

1. `transit_df` has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns, with each row representing one subway entrance or exit.

2. The data include the station, line, location, routes served, entrance type, entry, vending, and ADA status. I kept the columns needed for the analysis and changed `entry` from YES / NO to TRUE / FALSE.

3. The data are not tidy yet because the route information is spread across `route1` to `route11`.












