---
title: "p8105_hw2_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-10-06"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

## NYC Subway Data

```{r transit}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data - NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```
### Subway Data Description

1. `transit_df` has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns, with each row representing one subway entrance or exit.

2. The data include the station, line, location, routes served, entrance type, entry, vending, and ADA status. I kept the columns needed for the analysis and changed `entry` from YES / NO to TRUE / FALSE.

3. The data are not tidy yet because the route information is spread across `route1` to `route11`.


## Subway Stations

```{r station_questions}
n_stations = 
  transit_df |>
  distinct(line, station_name) |>
  nrow()

n_ada = 
  transit_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()

entry_prop = 
  transit_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

1. There are `r n_stations` distinct stations.

2. `r n_ada` stations are ADA compliant.

3. `r round(entry_prop * 100, 5)`% of station entrances / exits without vending allow entrance.

## Reformatting data 

```{r Reformatting}
transit_long = 
  transit_df |>
  mutate(
    across(route1:route11, as.character)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```


```{r}
n_a_stations = 
  transit_long |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

n_a_ada = 
  transit_long |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

1. The A train serves `r n_a_stations` distinct stations.

2. `r n_a_ada` of the stations served by the A train are ADA compliant.

# Problem 2

```{r}
library(readxl)
```


## Trash Wheel Data

```{r mr_trash}
mr_trash = 
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N742"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    sports_balls = as.integer(round(sports_balls)),
    wheel = "Mr. Trash Wheel"
  )

head(mr_trash, 5)
names(mr_trash)
dim(mr_trash)
```

### Mr. Trash Wheel Description

1. The Mr. Trash Wheel data have `r nrow(mr_trash)` observations after removing rows without dumpster data.

2. The variable names were cleaned using `clean_names()`.

3. `sports_balls` was rounded to the nearest whole number and converted to an integer.

## Professor Trash Wheel and Gwynnda

```{r}
prof_trash = 
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M139",
    na = "dive"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    wheel = "Professor Trash Wheel"
  )

gwynnda_trash = 
  readxl::read_excel(
    "./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = "A2:L394"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    wheel = "Gwynnda"
  )

head(prof_trash, 5)
head(gwynnda_trash, 5)

dim(prof_trash)
dim(gwynnda_trash)
```

The Professor Trash Wheel and Gwynnda data are cleaned in the same way, and `wheel` is added to identify the source of each observation.

## Combining Trash Wheel Data

```{r combine_trash}
trash_wheel = 
  bind_rows(
    mr_trash,
    prof_trash,
    gwynnda_trash
  ) |>
  select(wheel, everything())

head(trash_wheel, 5)
dim(trash_wheel)
```

```{r trash_variables}
trash_wheel |>
  select(
    wheel, dumpster, month, year,
    weight_tons, volume_cubic_yards,
    plastic_bottles, cigarette_butts
  ) |>
  head(10)
```

### Trash Wheel Data Description

1. Before combining the data, Mr. Trash Wheel has `r nrow(mr_trash)` observations, Professor Trash Wheel has `r nrow(prof_trash)` observations, and Gwynnda has `r nrow(gwynnda_trash)` observations. After using `bind_rows()`, the combined `trash_wheel` dataset has `r nrow(trash_wheel)` observations, with each row representing one dumpster collection.

2. The combined dataset includes key variables such as `wheel`, `dumpster`, `month`, `year`, `weight_tons`, `volume_cubic_yards`, `plastic_bottles`, and `cigarette_butts`. The `wheel` variable shows which Trash Wheel each observation comes from, while the other variables describe when the dumpster was collected and the amount or type of trash collected.

3. After combining the three datasets, observations from all three Trash Wheels are stored in the same dataframe. Variables with the same name are combined into the same columns, while variables that were not recorded for every Trash Wheel contain missing values for those observations.


## Trash Wheel Questions

For Professor Trash Wheel, we first filter the combined dataset to keep only observations from Professor Trash Wheel. We then focus on `weight_tons`, since this is the variable needed to answer the question about total trash collected. The first few observations are shown below to check the data used in the calculation.

```{r professor_data}
trash_wheel |>
  filter(wheel == "Professor Trash Wheel") |>
  select(wheel, dumpster, weight_tons) |>
  head(10)
```

The total weight is calculated by pulling `weight_tons` from the filtered data and summing all of the values.

```{r professor_weight}
prof_weight = 
  trash_wheel |>
  filter(wheel == "Professor Trash Wheel") |>
  pull(weight_tons) |>
  sum()

prof_weight
```

1. Professor Trash Wheel collected `r prof_weight` tons of trash in total.

For Gwynnda, we only need observations from June 2022. We therefore filter by `wheel`, `month`, and `year`, and then look at the `cigarette_butts` variable for those observations. The filtered rows are shown below before calculating the total.

```{r gwynnda_data}
trash_wheel |>
  filter(
    wheel == "Gwynnda",
    month == "June",
    year == 2022
  ) |>
  select(
    wheel, dumpster, month, year,
    cigarette_butts
  )
```

The total number of cigarette butts is calculated by pulling `cigarette_butts` from these June 2022 observations and summing the values.

```{r gwynnda_butts}
gwynnda_butts = 
  trash_wheel |>
  filter(
    wheel == "Gwynnda",
    month == "June",
    year == 2022
  ) |>
  pull(cigarette_butts) |>
  sum()

gwynnda_butts
```

2. Gwynnda collected `r gwynnda_butts` cigarette butts in June 2022.




# Problem 3

## Baseline Demographic Data

```{r}
baseline_all = 
  read_csv(
    "./data/data_mci/MCI_baseline.csv",
    skip = 1,
    na = "."
  ) |>
  janitor::clean_names() |>
  rename(
    age_baseline = current_age,
    age_mci = age_at_onset
  ) |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "Male",
      0 ~ "Female"
    ),
    sex = as.factor(sex),
    apoe4 = case_match(
      apoe4,
      1 ~ "Carrier",
      0 ~ "Non-carrier"
    ),
    apoe4 = as.factor(apoe4)
  )

head(baseline_all, 5)
names(baseline_all)
dim(baseline_all)
```

We check for participants who had already developed MCI by the time of the baseline visit and remove them because they do not meet the study inclusion criteria.

```{r}
baseline_excluded = 
  baseline_all |>
  filter(
    !is.na(age_mci),
    age_mci <= age_baseline
  )

baseline_excluded |>
  select(id, age_baseline, age_mci)

baseline_df = 
  baseline_all |>
  filter(
    is.na(age_mci) | age_mci > age_baseline
  )

head(baseline_df, 5)
dim(baseline_df)
```

### Baseline Data Description

1. The original baseline dataset has `r nrow(baseline_all)` participants. After removing `r nrow(baseline_excluded)` participants who had already developed MCI at baseline, `r nrow(baseline_df)` participants remain in the cleaned dataset.

2. The first row of the file contains descriptions rather than the actual variable names, so `skip = 1` was used when importing the data. `"."` was treated as a missing value for participants who did not develop MCI during follow-up.

3. The baseline dataset includes variables such as age, sex, education, APOE4 carrier status, and age of MCI onset. Since `sex` and `apoe4` were originally coded as 0 and 1, I changed them to `Female` / `Male` and `Non-carrier` / `Carrier` and stored them as factors.

## Baseline Questions

The main variables needed for these questions are baseline age, sex, APOE4 carrier status, and age of MCI onset. The first few observations are shown below.

```{r baseline_variables}
baseline_df |>
  select(
    id, age_baseline, sex,
    education, apoe4, age_mci
  ) |>
  head(10)
```

We first find the number of participants in the cleaned baseline dataset and how many developed MCI during follow-up.

```{r participant_questions}
n_recruited = 
  baseline_df |>
  nrow()

n_mci = 
  baseline_df |>
  filter(!is.na(age_mci)) |>
  nrow()

n_recruited
n_mci
```

1. After applying the study inclusion criteria, `r n_recruited` participants are included in the study. Of these participants, `r n_mci` developed MCI during follow-up.

The average baseline age is calculated using `age_baseline`.

```{r baseline_age}
avg_age = 
  baseline_df |>
  pull(age_baseline) |>
  mean()

avg_age
```

2. The average baseline age is `r round(avg_age, 2)` years.

To find the proportion of women who are APOE4 carriers, we first count the total number of women and then count how many of these women are APOE4 carriers.

```{r women_apoe}
n_women = 
  baseline_df |>
  filter(sex == "Female") |>
  nrow()

n_women_apoe = 
  baseline_df |>
  filter(
    sex == "Female",
    apoe4 == "Carrier"
  ) |>
  nrow()

women_apoe_prop = 
  n_women_apoe / n_women

n_women
n_women_apoe
women_apoe_prop
```

3. There are `r n_women` women in the study, and `r n_women_apoe` of them are APOE4 carriers. Therefore, `r round(women_apoe_prop * 100, 1)`% of women in the study are APOE4 carriers.









