---
title: "p8105_hw2_zc2959"
author: "Zhenghong Chen zc2959"
date: "2026-10-06"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1

## NYC Subway Data

```{r transit}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data - NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```
### Subway Data Description

1. `transit_df` has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns, with each row representing one subway entrance or exit.

2. The data include the station, line, location, routes served, entrance type, entry, vending, and ADA status. I kept the columns needed for the analysis and changed `entry` from YES / NO to TRUE / FALSE.

3. The data are not tidy yet because the route information is spread across `route1` to `route11`.


## Subway Stations

```{r station_questions}
n_stations = 
  transit_df |>
  distinct(line, station_name) |>
  nrow()

n_ada = 
  transit_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()

entry_prop = 
  transit_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```

1. There are `r n_stations` distinct stations.

2. `r n_ada` stations are ADA compliant.

3. `r round(entry_prop * 100, 5)`% of station entrances / exits without vending allow entrance.

## Reformatting data 

```{r Reformatting}
transit_long = 
  transit_df |>
  mutate(
    across(route1:route11, as.character)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```


```{r}
n_a_stations = 
  transit_long |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

n_a_ada = 
  transit_long |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

1. The A train serves `r n_a_stations` distinct stations.

2. `r n_a_ada` of the stations served by the A train are ADA compliant.





